#Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

#Goal
Goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise. This is the "classe" variable in the training set. We will use any of the other variables to predict with. The report describes how the model was built, how I used cross validation, what I think the expected out of sample error is, and why I made the choices you did. 

#Preprocessing
```{r echo=TRUE}
library(caret)
```

Load data and create training set with 60% of dataset.
```{r echo=TRUE}
df <- read.csv("pml-training.csv", na.strings=c("NA",""))
inTrain <- createDataPartition(df$classe, p=0.6, list=FALSE)
training <- df[inTrain,]
testing <- df[-inTrain,]

```
